---
title: "Discrete Ancestral State Reconstruction"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, include=FALSE}
#devtools::install_github("arborworkflows/aRbor")
library(aRbor)
library(geiger)
library(knitr)
library(pander)
get_char_ratio <- function(ace_arbor){
  trait_ratio = list()
  for(row in 1:(length(ace_arbor[[1]])/2) ){
    max = max(ace_arbor[[1]][row,])
    for (cell in 1:length(ace_arbor[[1]][row,] )){
      val = ace_arbor[[1]][row,cell]
      if (val == max){
        trait_ratio = c(trait_ratio, cell)
      }
    }
  }
  return(table(as.character(trait_ratio)))
}
input_table = function(ace_arbor){
  chartype = attributes(ace_arbor)$charType
  acetype = attributes(ace_arbor)$aceType
  numChar = length(attributes(ace_arbor)$charStates[[1]])
  lik =  attributes(ace_arbor[[1]])$fit$lnLik
  rate = attributes(ace_arbor[[1]])$fit$par.full

  ace_table <- t(data.frame("type of trait"=chartype, "asr method"=acetype, "number of traits"=numChar, "logLikelihood"=lik, 'transition rate'=rate))[,1]
  #names(ace_table) <- c("type", "value")
  return(ace_table)
}

output_table = function(ace_arbor){
  trait.ratio = get_char_ratio(ace_arbor)
  tipChar = as.matrix(table(attributes(ace_arbor)[[3]]$dat))
  
  ace_table <- data.frame('trait.ratio'=trait.ratio, "tipChar"=tipChar)
  names(ace_table) <- c("trait", "internal node frequency", 'tip frequency')
  rownames(ace_table) <- c()
  return(ace_table)
}
```

## Load your data and run the method

```{r, echo=TRUE, results='hide'}

tr  <- read.tree("/home/blubb/Documents/Workshops_PD/Ithaca_2018/example_data_ARBOR/Heliconia_all_dated_mod.phy")
trait <- read.csv('/home/blubb/Documents/Workshops_PD/Ithaca_2018/example_data_ARBOR/Heliconia_phylogeny_matrix.csv')
  ape::is.ultrametric(tr)
  if (  ape::is.ultrametric(tr) == FALSE){
    tr <- phytools::force.ultrametric(tr)
    print("Your tree was not ultrametric, I forced it to be")
  }
td = treeplyr::make.treedata(tr, trait)

ace_arbor = aRbor::aceArbor( select_(td, "Inflorescence"), charType='discrete', aceType = 'marginal')
```

## Summaries about your ASR

Besides calculating the ancestral states in the phylogeny, here are some summary information for discrete character.

Table 1 shows you your type of trait, input, reconstruction method, the log Likelihood and the transition rate between the traits.
```{r}
  ace_table = input_table(ace_arbor)

  kable(ace_table)
```

Table 2 sums up the result.
```{r}
ace_output = output_table(ace_arbor)

kable(ace_output)
```

## Some information might have been removed during analysis

Be aware that the function dropped some tips of your phylogeny:
```{r}
if (length(attributes(td)$dropped$dropped_from_data) >=1){
  kable(attributes(td)$dropped$dropped_from_data)
}else
   pander::pander("No data was dropped.")
```

It might has dropped information from your trait dara as well: 
```{r, eval = T}
if (length(attributes(td)$dropped$dropped_from_tree) >=1){
    kable(attributes(td)$dropped$dropped_from_tree)
}else
  pander::pander("No data was dropped.")
```
